---
title: "HW4-Data Science Math"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Using matrix operations, describe the solutions for the following family of equations:

* x + 2y - 3z = 5 
* 2x + y - 3z = 13 
* -x + y + 2z= -8

### import Libraries
```{r}
library(MASS)
```


### Creating the Matrix

```{r}
A<-matrix(c(1,2,-3,2,1,-3,-1,1,2),nrow=3,ncol=3,byrow=TRUE)

B<-matrix(c(5,13,-8),nrow=3,ncol=1,byrow=TRUE)
A;B
```

#### a.  Find the inverse of the above 3x3 (non-augmented) matrix.

```{r}
s<-fractions(solve(A))
s
```

#### b.  Solve for the solution using R.

```{r}
solve(A,B)
```

### c.  Modify the  3x3 matrix such that there exists only one non-zero variable in the solution set.

```{r}
A2=cbind(B,A[,2],A[,3])
A2
A2s<-solve(A2)%*% B
round(A2s,6)
```

## Question 2

Consider the matrix,q=matrix(c(3,1,4,4,3,3,2,3,2),nrow=3). Let b=c(1,4,5). Use Cramer's rule and R to determine the solution, x, to qx=b, if one exists.  Show all determinants.

```{r}
q=matrix(c(3,1,4,4,3,3,2,3,2),nrow=3)
b=c(1,4,5)
q;b
```

create a gramer's function

```{r}
gramers<-function(ma,mb,nround_off)
{
  d=det(q)
  dx=det(cbind(mb,ma[,2],ma[,3]))
  dy=det(cbind(ma[,1],mb,ma[,3]))
  dz=det(cbind(ma[,1],ma[,2],mb))
  
  x_val=dx/d
  y_val=dy/d
  z_val=dz/d
  
  result<-c(round(x_val,nround_off),round(y_val,nround_off),round(z_val,nround_off))
  return(result)
}
```

Solve
```{r}
gramers(q,b,6)
solve(q,b)
```

